Hidden Talents in Context: Multiverse analysis repository

### 2022-03-08 README.Rmd
CD revision 2, minor changes and updates

This repository contains data, code, and (some) output for a manuscript entitled "Hidden talents in context: Cognitive performance with abstract versus ecological stimuli among adversity-exposed youth" submitted to and under review at *Child Development*.

>>>>

This repository contains data, code, and (some) output for a manuscript entitled "Hidden talents in context: Cognitive performance with abstract versus ecological stimuli among adversity-exposed youth" accepted for publication at *Child Development*.

### 2022-03-08 manuscript/manuscript.Rmd
CD revision 2, minor changes and updates

\noindent Adversity-exposed youth tend to score lower on cognitive tests. However, the hidden talents approach proposes some abilities are enhanced by adversity, especially under ecologically relevant conditions. Sampling `r txt1_sam$non_arb$n$total` youth (*M~age~* = `r txt1_sam$non_arb$age$mean`, *SD~age~* = `r txt1_sam$non_arb$age$sd`; `r (txt1_sam$non_arb$sex$females/txt1_sam$non_arb$n$total * 100) %>% round(2)`% female; `r ((txt1_sam$non_arb$n$total-txt1_sam$non_arb$race$non_white)/txt1_sam$non_arb$n$total * 100) %>% round(2)`% White), we administered two versions of an attention shifting and working memory updating task---one abstract, one ecological. We measured environmental unpredictability, violence, and poverty, and tested adversity × task version interactions. There were no interactions for attention shifting. For working memory updating, youth exposed to violence and poverty scored lower than their peers with abstract stimuli but just as well with ecological stimuli. These results are striking compared to contemporary developmental science, which almost exclusively reports lowered performance among adversity-exposed youth.

@@

The hidden talents approach is rooted in an adaptation-based framework [see @ellis2017; @ellis2020; @frankenhuis2020; @frankenhuis2013 for reviews], which assumes development tailors skills and abilities in relation to the environment. Accordingly, exposure to adversity should shape skills and abilities well-suited for dealing with the challenges and opportunities associated with harsh and unpredictable environments. However, because specific dimensions of adversity (e.g., violence versus resource-scarcity) may pose unique challenges [@ellis2020; @frankenhuis2020; @mclaughlin2019; @sheridan2014], the hidden talents approach proposes that cognitive skills should match specific dimensions of adversity [@frankenhuis2020]. For instance, some evidence suggests that physically abused youth can detect angry facial expressions quickly and accurately [@gibb2009; @pollak2008; @pollak2009], people living in poverty have more accurate social perception [@bjornsdottir2017; @dietze2021; @piff2018], and that anxiously attached people can more accurately detect deception [@eindor2014; @shoda2013].

@@

Hidden talents may also emerge when testing *content and stimuli* are more ecologically relevant and concrete. Typical cognitive tests use abstract content that is largely detached from the real world. Such abstract materials may disadvantage youth whose childhood experiences involve less abstract, analytical problem solving [@ellis2020; @frankenhuis2013]. Cultural psychologists and anthropologists have demonstrated this fact across the world. For example, people from non-Western cultures tend to score lower than people in Western cultures on standard versions of cognitive tests, but this performance gap closes or disappears when more ecologically relevant measures are used [@banerjee2017; @grigorenko2004; @rogoff2017; @sternberg2001]. This work challenges the assumption that "baseline" performance is best captured by standard/abstract tests. In fact, they suggest that abstract stimuli, and more broadly standard testing conditions, are specific contexts that may not match one's lived experience [@doebel2020; @zuilkowski2016]. These ideas also raise an important question: do adversity-exposed youth score higher on tests with ecologically-relevant stimuli relative to those with abstract stimuli?

@@

Although research testing performance in ecologically relevant contexts/situations and using ecologically relevant materials have generated innovative ideas and findings, it remains unclear whether or how these ideas and findings can be practically applied. For example, narrow abilities, such as memorizing social dominance hierarchies or accurately detecting anger in faces, seem to have limited practical value in educational contexts. In addition, although executive functions and working memory are related to many educational outcomes, it is unclear how schools and teachers might leverage skills and abilities that specifically depend on inducing uncertainty. That seems impractical, at best, and unethical, at worst.

Here we address this issue by evaluating the role of ecologically relevant content in attention-shifting and working memory tasks. We define ecologically relevant content as testing material that resembles the local environment and therefore has more meaning and consequence. We used two criteria for stimulus selection. First, we selected stimuli that youth would encounter and interact with on a daily basis. Second, we selected stimuli believed to be more consequential or salient to adversity-exposed youth. For example, money is more consequential and perhaps salient to adversity-exposed youth than other neutral objects (a book, dishes, etc.). We target both within-subject and between-subject patterns of improvement. That is, improvements in performance relative to oneself (within-subjects -- one's performance on abstract compared to ecological task versions), relative to others (between-subjects -- high versus low adversity), or a combination, are theoretically and practically important. In the current study, we expected all youth would score higher on tasks with ecological stimuli. However, we expected this increase to be larger among youth living in adverse conditions compared to youth in supportive conditions.

To address our research questions, we sampled `r nrow(data2_non_arbitrary)` youth from a socioeconomically diverse population. Using interviews, self-reports, and school records, we measured exposure to environmental unpredictability, violence, and poverty. We used these information sources to arrive at a "best estimate" for each construct. Our a priori assumption was that each information source (e.g., interview, self-report, school provided information) would correlate within constructs, which was largely true (see Table \@ref(tab:table1)).

@@

```{r figure1, fig.width=6, dpi=600, fig.id = "figure1", fig.cap.style = "Image Caption", fig.cap='Schematic display of the attention shifting and working memory updating tasks: a) abstract attention shifting, b) ecological attention shifting, c) abstract working memory updating, and d) ecological working memory updating.'}
knitr::include_graphics("figure1.pdf")
```

@@

The current study collected data from `r txt1_sam$raw$n$total` youth (*M~age~* = `r txt1_sam$raw$age$mean`, *SD~age~* = `r txt1_sam$raw$age$sd`) from two sources across Salt Lake City, Utah. The first was a middle school and the second were 5 after-school clubs. The middle school sample (seventh and eighth graders, *N* = `r txt1_sam$raw$n$wjms`; `r txt1_sam$raw$sex$wjms$females` females) formed `r (txt1_sam$raw$n$wjms/txt1_sam$raw$n$total * 100) %>% round(2)`% of the total sample and is socioeconomically and ethnically diverse: `r data1_raw %>% filter(race == 1) %>% nrow()` youth identified as Asian, `r data1_raw %>% filter(race == 2) %>%nrow()` as Native American, `r data1_raw %>% filter(race == 3) %>% nrow()` as African American, `r data1_raw %>% filter(race == 4) %>% nrow()` as White, `r data1_raw %>% filter(race == 5) %>% nrow()` as Hispanic, `r data1_raw %>% filter(race == 6)  %>% nrow()` as Pacific Islander, `r data1_raw %>% filter(race == 7) %>% nrow()` with multiple racial identities, and `r data1_raw %>% filter(is.na(race)) %>% nrow()` were missing. According to school district records, `r (txt1_sam$raw$free_lunch$wjms/txt1_sam$raw$n$wjms * 100) %>% round(2)`% of students were receiving economic assistance based on family income and/or homelessness (i.e., free or reduced-price lunch or fee waivers). The club sample (*N* = `r txt1_sam$raw$n$bgc`; `r txt1_sam$raw$sex$bgc$female` females) contained a wider age range (*M~age~*= `r txt1_sam$raw$age$bgc$mean`, *SD~age~* = `r txt1_sam$raw$age$bgc$sd`) and were also ethnically diverse. `r (txt1_sam$raw$free_lunch$bgc/txt1_sam$raw$n$bgc * 100) %>% round(2)`% were receiving free or reduced-price lunch.

@@

Data collection involved six parts: consent/assent, demographics form, cognitive testing, structured interview, questionnaire, and debriefing. Consent was obtained from the primary parent or guardian before assessments and assent was obtained from the participant at the beginning of the session. On the demographics form, participants indicated if they were left-handed, their birthdate, and their racial and ethnic background.

@@

The interview contained two parts. The first part was about participants' home and family life, such as their family composition, the number of adults in the household, and occupation and education information about caregivers. Participants also reported on the number of romantic partners their caregivers had and if these partners lived with the participant. The second part was about residential mobility. Participants listed all the homes they lived in from birth to age 13. For each home, they reported at which ages they lived in the home. If they moved, they also reported any changes in their family structure (loss or gain of adults in the home), social network (e.g., a new school), and/or town or city.

@@

Accuracy on attention-shifting tasks is normally high, and was in the current sample for both versions (*M~standard~* = `r (txt2_dvs$abs_shifting_correct$mean/32 * 100) %>% round(2)`%, *SD~standard~* = `r (txt2_dvs$abs_shifting_correct$sd/32 * 100) %>% round(2)`%; *M~ecological~* = `r (txt2_dvs$eco_shifting_correct$mean/32 * 100) %>% round(2)`%, *SD~ecological~* = `r (txt2_dvs$eco_shifting_correct$sd/32 * 100) %>% round(2)`%).Thus, instead of accuracy, attention-shifting ability is measured by comparing reaction times on repeat trials to switch trials. Reaction times on repeat trials are normally faster than switch trials; average repeat trial reaction times on the abstract and ecological versions were `r txt2_dvs$abs_shifting_re_rt$mean` *ms* (*SD* = `r txt2_dvs$abs_shifting_re_rt$sd` *ms*) and `r txt2_dvs$eco_shifting_re_rt$mean` *ms* (*SD* = `r txt2_dvs$eco_shifting_re_rt$sd` *ms*), respectively. On switch trials, average reaction times were `r txt2_dvs$abs_shifting_sw_rt$mean` *ms* (*SD* = `r txt2_dvs$abs_shifting_sw_rt$sd` *ms*) and `r txt2_dvs$eco_shifting_sw_rt$mean` *ms* (*SD* = `r txt2_dvs$eco_shifting_sw_rt$sd` *ms*) on the standard and ecological versions.

To calculate attention-shifting scores, we first cleaned trial-level data [@friedman2008; @miyake2000]. We removed fast and slow reaction times and removed trials after incorrect responses (see [Attention Shifting Data Cleaning Procedure](https://osf.io/kvt5e/?view_only=92d55bd3b9834468acd3a7cef5c3febd) for details). We then subtracted the average reaction time on valid repeat trials from the average reaction time on valid switch trials to calculate switch cost scores (*M~switch~* - *M~repeat~*). The switch cost measures the relative delay in reaction time caused by switching between categorization rules. Average switch costs on the standard and ecological versions were `r txt2_dvs$abs_shifting_sc$mean` *ms* (*SD* = `r txt2_dvs$abs_shifting_sc$sd` *ms*) and `r txt2_dvs$eco_shifting_sc$mean` *ms* (*SD* = `r txt2_dvs$eco_shifting_sc$sd` *ms*), respectively (*r* = `r txt5_cor$shifting$r`, *n* = `r txt5_cor$shifting$n`). Higher switch costs mean longer reaction times on switch trials compared to repeat trials. Higher switch costs indicate less efficient attention shifting.

@@

Our measure of environmental unpredictability integrated information from the interview and questionnaire. The interview contained three variables of interest. The first measured the level of exposure to living in a biologically non-intact family from birth to age 13. Interviewers asked participants about their family composition and the adults that lived with them in the home. If a participant's family composition changed, for example, because of a divorce, interviewers recorded the ages when participants experienced the change and for how long it persisted. A family was considered non-intact if the participant did not live with both biological parents during any period from birth to age 13. Based on the earliest age of disruption, we indexed the onset of exposure to a non-intact family by creating a four-level variable from 0 to 3. A 0 indicated the participant lived in an intact family (lived with both biological parents) from birth until age 13 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 0) %>% sum(na.rm=T)`); a 1 indicated that a disruption occurred after the age of 5 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 1) %>% sum(na.rm=T)`); a 2 indicated that a disruption occurred before age 5 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 2) %>% sum(na.rm=T)`); and a 3 indicated the participant was born into a non-intact family (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 3) %>% sum(na.rm=T)`). `r is.na(data2_non_arbitrary$unp_interview_disrupted) %>% sum(na.rm=T) %>% english::as.english() %>% str_to_title()` participants had missing information.

The second interview variable tabulated the number of different parental figures (defined as spouses or boyfriends/girlfriends of the participant's mother and father) who lived in the same home with the participant. Trained interviewers tabulated the total number of co-residing parental figures, excluding the participant's biological mother and father (*M* = `r txt3_ivs$unp_interview_figures$mean`, *SD* = `r txt3_ivs$unp_interview_figures$sd`).

@@

The questionnaire included a measure of perceived childhood unpredictability [@mittal2015; @young2018; @szepsenwol2015]. Participants responded to six items that assessed exposure to unpredictability up to age 13: 1) "Things were pretty calm and stable in my house" (reversed); 2) "People moved in and out of my house a lot"; 3) "I had a hard time knowing what my parent(s) or other people in my house were going to say or do from day-to-day"; 4) "My parent(s) argued or fought a lot with each other or with other people"; 5) "My parent(s) changed jobs a lot"; 6) "My parental situation changed a lot (for example, divorce or separation of parents, parents starting new romantic relationships, parents leaving the home)". Participants rated each item on a scale from 1 (never true) to 5 (very often true). Responses to these items were averaged to create a self-reported perceived childhood unpredictability measure (*M* = `r txt3_ivs$unp_perceived$mean`, *SD* = `r txt3_ivs$unp_perceived$sd`, $\alpha$ = `r txt4_rel$unp`). Higher scores indicated more perceived exposure to unpredictable environments.

@@

Participants also reported their perceived level of resource scarcity. They rated seven items from 1 (never true) to 5 (very often true): 1) "Your family had enough money to afford the kind of home you all needed"; "Your family had enough money to afford the kind of clothing you all needed"; 3) "Your family had enough money to afford the kind of food that you all needed"; 4) "Your family had enough money to afford the kind of medical care that you all needed"; 5) "I felt well-off (rich, wealthy) compared to other kids in my school"; 6) "I felt well-off (rich, wealthy) compared to other kids in my neighborhood"; and 7) "Your family struggled to make ends meet (get by financially)". The first six items were reversed, and all items were averaged to create a perceived resource scarcity measure (*M* = `r txt3_ivs$ses_perceived$mean`, *SD* = `r txt3_ivs$ses_perceived$sd`, $\alpha$ = `r txt4_rel$ses`). Higher scores indicated more perceived resource scarcity.

To create a poverty exposure composite, we combined parent education and occupational prestige (reversed), school/club reported economic disadvantage, and perceived resource scarcity. First, we standardized and averaged parent education and occupation (*r* = `r txt5_cor$edu_int$r`). Then, we recoded school-provided economic codes to be equal to half of a standard deviation above and below the mean (in terms of z-scores). Specifically, those receiving free or reduced-price lunch were coded as .5, and those who were not were coded as -.5. To create a final composite, we averaged parent education/occupation (standardized), perceived resource scarcity (standardized), and the recoded school/club economic disadvantage. Higher values indicated more exposure to poverty.

@@

We included two control variables: age and interviewer-rated distractions during testing. We included these because older youth tend to score higher on cognitive tests than younger youth [@best2010; @best2011; @huizinga2006], and disruptions during the tasks could negatively impact performance. Interviewers completed quantitative ratings of the testing environment including 1) the level of noise; 2) the level of chaos; 3) how often there were interruptions; and 4) how often the participant appeared distracted. Interviewers rated each item on a scale from 1 (not at all) to 7 (much more than average). Items were averaged to create a composite (*M* = `r txt3_ivs$test_envr$mean`, *SD* = `r txt3_ivs$test_envr$sd`), where higher scores indicate more distractions.

@@

We used a multiverse analysis approach for all analyses to systematically evaluate the robustness (or sensitivity) of analyses across all arbitrary data processing decisions [see @simonsohn2020; @steegen2016 for details]. We identified six arbitrary data processing decisions. First, we conducted analyses with and without the after-school club sample. We did this because our procedure was slightly different across the two samples and the testing environment was more disruptive in the clubs compared with the middle school. Second, we conducted analyses with and without participants with minor distractions during testing. Third, we analyzed the data with and without participants who missed one trap question (those who missed more than 1 were already excluded). Fourth, we analyzed data with and without participants having any amount of special education (participants receiving more \> 60 minutes were already excluded). Fifth, we analyzed data with and without participants who performed at or below chance levels of accuracy on the attention-shifting task (see [Attention Shifting Data Cleaning Procedure](https://osf.io/kvt5e/?view_only=92d55bd3b9834468acd3a7cef5c3febd) for more information).

@@

We preregistered all primary analyses on the Open Science Framework (OSF; see the [osf project link](https://osf.io/6r95z/?view_only=92d55bd3b9834468acd3a7cef5c3febd)), including an [Updated Primary Analysis Plan](https://osf.io/4vsnz/?view_only=92d55bd3b9834468acd3a7cef5c3febd), which included more details. We conducted secondary analyses after our primary analyses (see [Secondary Analyses](#secondary) below). Before conducting these analyses, we uploaded secondary analysis plans (see the [First Secondary Analysis Plan](https://osf.io/7fu35/?view_only=92d55bd3b9834468acd3a7cef5c3febd) and the [Second Plan](https://osf.io/wcauf/?view_only=92d55bd3b9834468acd3a7cef5c3febd)), explaining the goals and purpose of each secondary analysis. A Red Team Critic [@redteammarket2021] successfully reproduced and verified (see this [Red Team Report](https://osf.io/xv94z/?view_only=92d55bd3b9834468acd3a7cef5c3febd)) all analyses prior to submission.

@@

```{r figure2, fig.id = "figure2", fig.cap.style = "Image Caption", fig.width=6.5, fig.height=7.25, dpi=600, fig.cap = "Visualization of multiverse attention shifting results. Specifically, a) visualizes the multiverse task version × adversity interaction on abstract and ecological task versions across high (+ 1 SD) and low (- 1 SD) adversity exposure (y-axis was reversed so that higher values = faster shifting), b) plots p-curves associated with each interaction term, c) plots sorted interaction B-coefficients across each arbitrary decision, d) plots the sample sizes for each effect, and e) is a specification grid indicating the data processing decisions associated with each effect."}
fig2
```

>>>>

\noindent Adversity-exposed youth tend to score lower on cognitive tests. However, the hidden talents approach proposes some abilities are enhanced by adversity, especially under ecologically relevant conditions. Two versions of an attention shifting and working memory updating task---one abstract, one ecological---were administered to `r txt1_sam$non_arb$n$total` youth (*M~age~* = `r txt1_sam$non_arb$age$mean`, *SD~age~* = `r txt1_sam$non_arb$age$sd`; `r (txt1_sam$non_arb$sex$females/txt1_sam$non_arb$n$total * 100) %>% round(2)`% female; `r ((txt1_sam$non_arb$n$total-txt1_sam$non_arb$race$non_white)/txt1_sam$non_arb$n$total * 100) %>% round(2)`% White). Measures of environmental unpredictability, violence, and poverty were collected to tested adversity × task version interactions. There were no interactions for attention shifting. For working memory updating, youth exposed to violence and poverty scored lower than their peers with abstract stimuli but equally well with ecological stimuli. These results are striking compared to contemporary developmental science, which often reports lowered performance among adversity-exposed youth.

@@

The hidden talents approach is rooted in an adaptation-based framework [see @ellis2017; @ellis2020; @frankenhuis2020; @frankenhuis2013 for reviews], which assumes development tailors skills and abilities in relation to the environment. Accordingly, exposure to adversity should shape skills and abilities well-suited for dealing with the challenges and opportunities associated with harsh and unpredictable environments. However, because specific dimensions of adversity (e.g., violence versus resource-scarcity) may pose unique challenges [@ellis2020; @frankenhuis2020; @mclaughlin2019; @sheridan2014], the hidden talents approach proposes that cognitive skills should match specific dimensions of adversity [@frankenhuis2020]. For instance, some evidence suggests that physically abused youth can detect angry facial expressions quickly and accurately [@gibb2009; @pollak2008; @pollak2009], lower perceived social class or status predicts more accurate social perceptions [@bjornsdottir2017; @dietze2021; @piff2018], and that anxiously attached people can more accurately detect deception [@eindor2014; @shoda2013].

@@

Hidden talents may also emerge when testing *content and stimuli* are more ecologically relevant and concrete. Typical cognitive tests use abstract content that is largely detached from the real world. Such abstract materials may disadvantage youth whose childhood experiences involve less abstract, analytical problem solving [@ellis2020; @frankenhuis2013]. Cultural psychologists and anthropologists have demonstrated this fact across the world. For example, people from non-Western cultures tend to score lower than people in Western cultures on standard versions of cognitive tests, but this performance gap closes or disappears when more ecologically relevant measures are used [@banerjee2017; @grigorenko2004; @rogoff2017; @sternberg2001]. This work challenges the assumption that "baseline" performance is best captured by standard, abstract tests. In fact, they suggest that abstract stimuli, and more broadly standard testing conditions, are specific contexts that may not match one's lived experience [@doebel2020; @miller-cotto2021; @nketia2021 ; @zuilkowski2016]. These ideas also raise an important question: do adversity-exposed youth score higher on tests with ecologically-relevant stimuli relative to those with abstract stimuli?

@@

Although research testing performance in ecologically relevant contexts or situations and using ecologically relevant materials have generated innovative ideas and findings, it remains unclear whether or how these ideas and findings can be practically applied. For example, narrow abilities, such as memorizing social dominance hierarchies or accurately detecting anger in faces, seem to have limited practical value in educational contexts. In addition, although executive functions and working memory are related to many educational outcomes, it is unclear how schools and teachers might leverage skills and abilities that specifically depend on inducing uncertainty. That seems impractical, at best, and unethical, at worst.

Here we address this issue by evaluating the role of ecologically relevant content in attention-shifting and working memory tasks. We define ecologically relevant content as testing material that resembles the local environment and therefore has more meaning and consequence. We used two criteria for stimulus selection. First, we selected stimuli that youth would encounter and interact with on a daily basis. Second, we selected stimuli believed to be more consequential or salient to adversity-exposed youth. For example, money is more consequential to poverty-exposed youth and an angry face more salient to violence-exposed youth than neutral objects (a book, dishes, etc.). Although some stimuli are likely to be more relevant to one adversity dimension than another, different adversity exposures tend to co-occur [@jacobs2012; @smith2021]. Thus, we selected stimuli that are broadly relevant in the context of correlated adversity exposures.

We target both within-subject and between-subject patterns of improvement. That is, improvements in performance relative to oneself (within-subjects -- one's performance on abstract compared to ecological task versions), relative to others (between-subjects -- high versus low adversity), or a combination, are theoretically and practically important. In the current study, we expected all youth would score higher on tasks with ecological stimuli. However, we expected this increase to be larger among youth living in adverse conditions compared to youth in supportive conditions.

To address our research questions, we sampled `r nrow(data2_non_arbitrary)` youth from a socioeconomically diverse population to ensure sufficient variation in adversity-exposure. Using interviews, self-reports, and school records, we measured exposure to environmental unpredictability, violence, and poverty. We used these information sources to arrive at a "best estimate" for each construct. Our a priori assumption was that each information source (e.g., interview, self-report, school provided information) would correlate within constructs, which was largely true (see Table \@ref(tab:table1)).

@@

```{r figure1, fig.width=6, dpi=600, fig.id = "figure1", fig.cap.style = "Image Caption", fig.cap='Schematic display of the attention shifting and working memory updating tasks: a) abstract attention shifting, b) ecological attention shifting, c) abstract working memory updating, and d) ecological working memory updating.'}
knitr::include_graphics("figures/figure1.pdf")
```

@@

The current study collected data from `r txt1_sam$raw$n$total` youth (*M~age~* = `r txt1_sam$raw$age$mean`, *SD~age~* = `r txt1_sam$raw$age$sd`) from two sources across Salt Lake City, Utah, USA starting in the spring of 2018 and ending in early 2020. The first was a middle school and the second were 5 after-school clubs. The middle school sample (seventh and eighth graders, *N* = `r txt1_sam$raw$n$wjms`; `r txt1_sam$raw$sex$wjms$females` females) formed `r (txt1_sam$raw$n$wjms/txt1_sam$raw$n$total * 100) %>% round(2)`% of the total sample and is socioeconomically and ethnically diverse: `r data1_raw %>% filter(race == 1) %>% nrow()` youth identified as Asian, `r data1_raw %>% filter(race == 2) %>%nrow()` as Native American, `r data1_raw %>% filter(race == 3) %>% nrow()` as African American, `r data1_raw %>% filter(race == 4) %>% nrow()` as White, `r data1_raw %>% filter(race == 5) %>% nrow()` as Hispanic, `r data1_raw %>% filter(race == 6)  %>% nrow()` as Pacific Islander, `r data1_raw %>% filter(race == 7) %>% nrow()` with multiple racial identities, and `r data1_raw %>% filter(is.na(race)) %>% nrow()` were missing. According to school district records, `r (txt1_sam$raw$free_lunch$wjms/txt1_sam$raw$n$wjms * 100) %>% round(2)`% of students were receiving economic assistance based on family income and homelessness (i.e., free or reduced-price lunch or fee waivers). The club sample (*N* = `r txt1_sam$raw$n$bgc`; `r txt1_sam$raw$sex$bgc$female` females) contained a wider age range (*M~age~*= `r txt1_sam$raw$age$bgc$mean`, *SD~age~* = `r txt1_sam$raw$age$bgc$sd`) and were also ethnically diverse. `r (txt1_sam$raw$free_lunch$bgc/txt1_sam$raw$n$bgc * 100) %>% round(2)`% were receiving free or reduced-price lunch.

@@

Data collection involved six parts: consent and assent, demographics form, cognitive testing, structured interview, questionnaire, and debriefing. Consent was obtained from the primary parent or guardian before assessments and assent was obtained from the participant at the beginning of the session. On the demographics form, participants indicated if they were left-handed, their birthdate, and their racial and ethnic background.

@@

The interview contained two parts. The first part was about participants' home and family life, such as their family composition, the number of adults in the household, and occupation and education information about caregivers. Participants also reported on the number of romantic partners their caregivers had and if these partners lived with the participant. The second part was about residential mobility. Participants listed all the homes they lived in from birth to age 13. For each home, they reported at which ages they lived in the home. If they moved, they also reported any changes in their family structure (loss or gain of adults in the home), social network (e.g., a new school), and town or city.

@@

Accuracy on attention-shifting tasks is normally high, and was in the current sample for both versions (*M~standard~* = `r (txt2_dvs$abs_shifting_correct$mean/32 * 100) %>% round(2)`%, *SD~standard~* = `r (txt2_dvs$abs_shifting_correct$sd/32 * 100) %>% round(2)`%; *M~ecological~* = `r (txt2_dvs$eco_shifting_correct$mean/32 * 100) %>% round(2)`%, *SD~ecological~* = `r (txt2_dvs$eco_shifting_correct$sd/32 * 100) %>% round(2)`%). Thus, instead of accuracy, attention-shifting ability is measured by comparing reaction times on repeat trials to switch trials. Reaction times on repeat trials are normally faster than switch trials; average repeat trial reaction times on the abstract and ecological versions were `r txt2_dvs$abs_shifting_re_rt$mean` *ms* (*SD* = `r txt2_dvs$abs_shifting_re_rt$sd` *ms*) and `r txt2_dvs$eco_shifting_re_rt$mean` *ms* (*SD* = `r txt2_dvs$eco_shifting_re_rt$sd` *ms*), respectively. On switch trials, average reaction times were `r txt2_dvs$abs_shifting_sw_rt$mean` *ms* (*SD* = `r txt2_dvs$abs_shifting_sw_rt$sd` *ms*) and `r txt2_dvs$eco_shifting_sw_rt$mean` *ms* (*SD* = `r txt2_dvs$eco_shifting_sw_rt$sd` *ms*) on the standard and ecological versions.

To calculate attention-shifting scores, we first cleaned trial-level data [@friedman2008; @miyake2000]. We removed fast and slow reaction times and removed trials after incorrect responses (see <https://osf.io/kvt5e/> for details). We then subtracted the average reaction time on valid repeat trials from the average reaction time on valid switch trials to calculate switch cost scores (*M~switch~* - *M~repeat~*). The switch cost measures the relative delay in reaction time caused by switching between categorization rules. Average switch costs on the standard and ecological versions were `r txt2_dvs$abs_shifting_sc$mean` *ms* (*SD* = `r txt2_dvs$abs_shifting_sc$sd` *ms*) and `r txt2_dvs$eco_shifting_sc$mean` *ms* (*SD* = `r txt2_dvs$eco_shifting_sc$sd` *ms*), respectively (*r* = `r txt5_cor$shifting$r`, *n* = `r txt5_cor$shifting$n`). Higher switch costs mean longer reaction times on switch trials compared to repeat trials. Higher switch costs indicate less efficient attention shifting.

@@

Environmental unpredictability is difficult to measure and different definitions exist alongside each other [@young2020]. Despite this, we aligned our measurement approach with previous literature in two ways. First, we measured caregiver and residential instability, which are both thought to be proxies to environmental unpredictability, using interview-based measures [@ellis2009]. Second, we measured the extent to which youth perceive their family context as unpredictable using a questionnaire [@young2018]. Our goal was to aggregate these two information sources to arrive at a best estimate of environmental unpredictability.

The interview contained three variables of interest. The first measured the level of exposure to living in a biologically non-intact family from birth to age 13. Interviewers asked participants about their family composition and the adults that lived with them in the home. If a participant's family composition changed, for example, because of a divorce, interviewers recorded the ages when participants experienced the change and for how long it persisted. A family was considered non-intact if the participant did not live with both biological parents during any period from birth to age 13. Based on the earliest age of disruption, we indexed the onset of exposure to a non-intact family by creating a four-level variable from 0 to 3. A 0 indicated the participant lived in an intact family (lived with both biological parents) from birth until age 13 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 0) %>% sum(na.rm=T)`); a 1 indicated that a disruption occurred after the age of 5 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 1) %>% sum(na.rm=T)`); a 2 indicated that a disruption occurred before age 5 (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 2) %>% sum(na.rm=T)`); and a 3 indicated the participant was born into a non-intact family (*n* = `r (data2_non_arbitrary$unp_interview_disrupted == 3) %>% sum(na.rm=T)`). `r is.na(data2_non_arbitrary$unp_interview_disrupted) %>% sum(na.rm=T) %>% english::as.english() %>% str_to_title()` participants had missing information.

The second interview variable tabulated the number of different parental figures (defined as spouses or boyfriends or girlfriends of the participant's mother and father) who lived in the same home with the participant. Trained interviewers tabulated the total number of co-residing parental figures, excluding the participant's biological mother and father (*M* = `r txt3_ivs$unp_interview_figures$mean`, *SD* = `r txt3_ivs$unp_interview_figures$sd`).

@@

The questionnaire included a measure of perceived childhood unpredictability [@mittal2015; @young2018]. Participants responded to six items that assessed exposure to unpredictability up to age 13: 1) "Things were pretty calm and stable in my house" (reversed); 2) "People moved in and out of my house a lot"; 3) "I had a hard time knowing what my parent(s) or other people in my house were going to say or do from day-to-day"; 4) "My parent(s) argued or fought a lot with each other or with other people"; 5) "My parent(s) changed jobs a lot"; 6) "My parental situation changed a lot (for example, divorce or separation of parents, parents starting new romantic relationships, parents leaving the home)". Participants rated each item on a scale from 1 (never true) to 5 (very often true). Responses to these items were averaged to create a self-reported perceived childhood unpredictability measure (*M* = `r txt3_ivs$unp_perceived$mean`, *SD* = `r txt3_ivs$unp_perceived$sd`, $\alpha$ = `r txt4_rel$unp`). Higher scores indicated more perceived exposure to unpredictable environments.

@@

Participants also reported their perceived level of resource scarcity. They rated seven items from 1 (never true) to 5 (very often true): 1) "Your family had enough money to afford the kind of home you all needed"; 2) "Your family had enough money to afford the kind of clothing you all needed"; 3) "Your family had enough money to afford the kind of food that you all needed"; 4) "Your family had enough money to afford the kind of medical care that you all needed"; 5) "I felt well-off (rich, wealthy) compared to other kids in my school"; 6) "I felt well-off (rich, wealthy) compared to other kids in my neighborhood"; and 7) "Your family struggled to make ends meet (get by financially)". The first six items were reversed, and all items were averaged to create a perceived resource scarcity measure (*M* = `r txt3_ivs$ses_perceived$mean`, *SD* = `r txt3_ivs$ses_perceived$sd`, $\alpha$ = `r txt4_rel$ses`). Higher scores indicated more perceived resource scarcity.

To create a poverty exposure composite, we combined parent education and occupational prestige (reversed), school or club reported economic disadvantage, and perceived resource scarcity. First, we standardized and averaged parent education and occupation (*r* = `r txt5_cor$edu_int$r`). Then, we recoded school-provided economic codes to be equal to half of a standard deviation above and below the mean (in terms of z-scores). Specifically, those receiving free or reduced-price lunch were coded as .5, and those who were not were coded as -.5. To create a final composite, we averaged parent education and occupation (standardized), perceived resource scarcity (standardized), and the recoded school or club economic disadvantage. Higher values indicated more exposure to poverty.

@@

We included two control variables: age and interviewer-rated distractions during testing. We included these because older youth tend to score higher on cognitive tests than younger youth [@best2011], and disruptions during the tasks could negatively impact performance. Interviewers completed quantitative ratings of the testing environment including 1) the level of noise; 2) the level of chaos; 3) how often there were interruptions; and 4) how often the participant appeared distracted. Interviewers rated each item on a scale from 1 (not at all) to 7 (much more than average). Items were averaged to create a composite (*M* = `r txt3_ivs$test_envr$mean`, *SD* = `r txt3_ivs$test_envr$sd`), where higher scores indicate more distractions.

@@

We used a multiverse analysis approach for all analyses to systematically evaluate the robustness (or sensitivity) of analyses across all arbitrary data processing decisions [see @simonsohn2020; @steegen2016 for details]. We identified six arbitrary data processing decisions. First, we conducted analyses with and without the after-school club sample. We did this because our procedure was slightly different across the two samples and the testing environment was more disruptive in the clubs compared with the middle school. Second, we conducted analyses with and without participants with minor distractions during testing. Third, we analyzed the data with and without participants who missed one trap question (those who missed more than 1 were already excluded). Fourth, we analyzed data with and without participants having any amount of special education (participants receiving more \> 60 minutes were already excluded). Fifth, we analyzed data with and without participants who performed at or below chance levels of accuracy on the attention-shifting task (see <https://osf.io/kvt5e/> for more information).

@@

We preregistered all primary analyses on the Open Science Framework (OSF; see <https://osf.io/6r95z/>, including an updated primary analysis plan (see <https://osf.io/4vsnz/>), which included more details. We conducted secondary analyses after our primary analyses (see [Secondary Analyses](#secondary) below). Before conducting these analyses, we uploaded secondary analysis plans (see <https://osf.io/7fu35/> for the first plan and <https://osf.io/wcauf/> for the second plan), explaining the goals and purpose of each secondary analysis. A Red Team Critic [@redteammarket2021] successfully reproduced and verified (see <https://osf.io/xv94z/> for a report) all analyses prior to submission. All associated data and code are included in our preregistration repository on GitHub (<https://github.com/ethan-young/hidden-talents-multiverse>) and the OSF (<https://osf.io/6r95z/>), including source code to reproduce the manuscript itself.

@@

```{r figure2, fig.path="figures/", dev="pdf", fig.id = "figure2", fig.cap.style = "Image Caption", fig.width=6.5, fig.height=7.25, dpi=600, fig.cap = "Visualization of multiverse attention shifting results. Specifically, a) visualizes the multiverse task version × adversity interaction on abstract and ecological task versions across high (+ 1 SD) and low (- 1 SD) adversity exposure (y-axis was reversed so that higher values = faster shifting), b) plots p-curves associated with each interaction term, c) plots sorted interaction B-coefficients across each arbitrary decision, d) plots the sample sizes for each effect, and e) is a specification grid indicating the data processing decisions associated with each effect."}
fig2
```
